---
title: "Kinases"
output: html_notebook
---


```{r setup}
library(tidyverse)
library(folderfun)
library(readxl)
library(ggtext)
library(survival)
library(ggfortify)
library(survminer)
library(glue)

theme_set(theme_minimal())

setff("DataFolder", file.path(Sys.getenv("DiezmannLab"), "kinases", "data_files"))
setff("FigFolder", file.path(Sys.getenv("DiezmannLab"), "kinases", "Figures"))
dir.create(ffFigFolder(), showWarnings = FALSE)

number_groups <- function(data, col) { data %>% mutate(bio_rep = group_indices_(., .dots=col)) }

genotypes=data.frame(YSD=c("89", "577", "622", "628", "634",
                           "591", "667", "911", "687",
                           "883", "885", "961", "979"), 
                     genotype=c("WT", "cka1∆/∆", "cka2∆/∆", "ckb1∆/∆", "ckb2∆/∆",
                                "fun31∆/∆", "thi20∆/∆", "orf19.792", "rim15∆/∆",
                                "hog1∆/∆", "pbs2∆/∆", "mkk2∆/∆", "pkc1∆/∆"),
                     category=c("WT", rep("CK2", 4), rep("FSK", 4), rep("Conserved", 4)))

list_bio_reps <- function(data) {
  data %>% nest() %>%
  mutate(data=modify(data, number_groups, "date")) %>%
  unnest(cols = c(data)) %>%
  mutate(bio_rep=ifelse(genotype=='WT', NA_integer_, bio_rep))
}

get_genotypes <- function(data, metadata, genotype_order){
  genotype_order <- ifelse(genotype_order=='WT', genotype_order, glue("*{genotype_order}*"))
  data %>%  filter(! genotype %in% c('orf', 'orf19')) %>%
    mutate(genotype=tolower(genotype)) %>%
    left_join(filter(metadata, column_name=='genotype'),
            by=c("genotype" = "code")) %>% 
  select(-column_name) %>%
  rename(code=genotype, genotype=meaning) %>%
  mutate(genotype = factor(ifelse(genotype=='WT', genotype, glue("*{genotype}*")),
                           levels=genotype_order))
}

make_boxplot <- function(data, x, y, colour, facet_x, facet_y, xlab="", ylab="") {
    ggplot(data, aes(x={{x}}, y={{y}})) + 
    geom_boxplot(colour='grey50', fill='black', alpha=.1, outlier.alpha=0) +
    geom_point(aes(colour={{colour}}), size=2, alpha=.25, stroke=0) +
    facet_grid(rows=enquos(facet_y), cols=enquos(facet_x)) +
    scale_colour_brewer(palette="Set1", name="Biological Rep", na.value='grey50') +
    xlab({{xlab}}) +
    ylab({{ylab}}) +
    theme(strip.text.y = element_markdown(angle=0),
          strip.text.x = element_markdown(),
          panel.grid.major=element_line(size=.25),
          legend.position = "bottom",
          panel.grid.minor.x=element_blank()) +
    guides(colour = guide_legend(override.aes = list(alpha = 1)))
}

```

# antifungals
plot as heat map
normalise to the no gda no drug control for each strain
```{r antifungals}
antifungals <- read_excel(ffDataFolder("antifungals.xlsx"))
antifungals_metadata <- read_excel(ffDataFolder("antifungals.xlsx"), sheet="code")

control <- antifungals %>% filter(gda==0 & concentration==0) %>% select(-gda, -concentration, control_od = od)
(antifungals_heatmap <- 
antifungals %>% 
  filter(drug == 'fluconazole') %>%
  left_join(control) %>% 
  mutate(normalised_od = od/control_od) %>% 
  left_join(filter(antifungals_metadata, column_name=='genotype'),
            by=c("genotype" = "code")) %>% 
  select(-column_name) %>%
  rename(code=genotype, genotype=meaning) %>%
  mutate(concentration=round(concentration, 4),
         gda=glue("{gda} µg/ml GDA"),
         genotype = factor(ifelse(genotype=='WT', genotype, glue("*{genotype}*")),
                           levels=rev(ifelse(genotypes$genotype=='WT', genotypes$genotype, glue("*{genotypes$genotype}*"))))) %>%
  filter(genotype != "orf19") %>% 
  ggplot(aes(x=as.factor(concentration), 
             y=genotype, 
             fill=normalised_od)) +
    geom_tile() +
    facet_grid(. ~gda, scales="free") +
    #facet_grid(gda ~ date, scales="free") +
    scale_fill_viridis_c(name="Normalised OD") +
    xlab("Concentration (µg/ml)") +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5),
          axis.text.y=element_markdown(),
          axis.title.y=element_blank())) 

ggsave(ffFigFolder("antifungals_heatmap.png"), antifungals_heatmap)
```
# bf_biomass
plot as scatter with mean and deviation

```{r bf_biomass}
bf_biomass <- read_excel(ffDataFolder("bf_biomass.xlsx"))
bf_biomass_metadata <- read_excel(ffDataFolder("bf_biomass.xlsx"), sheet="code")

(bf_biomass_plot <-
  bf_biomass %>% 
  get_genotypes(bf_biomass_metadata, genotypes$genotype) %>%
  make_boxplot(x=mass, y=as.factor(gda), colour=NULL, NULL, genotype, xlab="mass", ylab="GDA (µg/ml)"))


ggsave(ffFigFolder("bf_biomass.png"), bf_biomass_plot)
```

# colony morphology
sectored pie charts for different replicates

```{r colony_morphology}
colony_morphology <- read_excel(ffDataFolder("colony_morphology.xlsx"))
colony_morphology_metadata <- read_excel(ffDataFolder("colony_morphology.xlsx"), sheet="code")

(colony_morphology_plot <-
colony_morphology %>% 
  get_genotypes(colony_morphology_metadata, genotypes$genotype) %>% 
  gather("rep", "score", starts_with("R")) %>%
  mutate(media=relevel(factor(media), ref='ypd')) %>% 
  ggplot(aes(x="", y="", fill=factor(score))) +
    geom_col() +
    coord_polar("y") +
    #scale_fill_viridis_d(name="Morphology score") +
    scale_fill_brewer(palette = "Set1") +
    facet_grid(genotype ~ media) +
    theme(strip.text.y =element_markdown(angle=0),
          strip.text.x =element_text(angle=60),
          axis.title = element_blank(),
          panel.grid.major=element_blank()))


ggsave(ffFigFolder("colony_morphology.png"), colony_morphology_plot)
```
# heat shock
5 technical replicates (r1-r5) and three. Biological replicates
plot as scatter with mean and deviation
```{r heat_shock}
heat_shock <- read_excel(ffDataFolder("heat_shock.xlsx"))
heat_shock_metadata <- read_excel(ffDataFolder("heat_shock.xlsx"), sheet="code")

(heat_shock_plot <-
heat_shock %>% 
  get_genotypes(heat_shock_metadata, genotypes$genotype) %>% 
  mutate(temperature=glue("{temperature}ºC")) %>%  
  gather("rep", "cfu", starts_with("r")) %>% 
  group_by(genotype, gda, temperature) %>% 
  list_bio_reps() %>%
  make_boxplot(x=cfu, y=as.factor(gda), colour=factor(bio_rep), temperature, genotype, xlab="CFU", ylab="GDA (µg/ml)") +
    scale_x_continuous(breaks = c(25,50,100,200,400,800)) +
    coord_trans('log10')
)
  
ggsave(ffFigFolder("heat_shock.png"), heat_shock_plot, height=6)
```

# oxidative stress
plot as scatter with mean and deviation
here distinguish between treated with and without H2O2

```{r oxidative_stress}
oxidative_stress <- read_excel(ffDataFolder("oxidative_stress.xlsx"))
oxidative_stress_metadata <- read_excel(ffDataFolder("oxidative_stress.xlsx"), sheet="code")


(oxidative_stress_plot <-
oxidative_stress %>% 
  get_genotypes(oxidative_stress_metadata, genotypes$genotype) %>%
  mutate(h2o2 = glue("{h2o2} mM H<sub>2</sub>O<sub>2</sub>")) %>%  
  gather("rep", "cfu", starts_with("r")) %>%
  group_by(genotype, gda, h2o2) %>% 
  list_bio_reps() %>%
  mutate(bio_rep=ifelse(genotype=='WT', NA_integer_, bio_rep)) %>%
  make_boxplot(x=cfu, y=as.factor(gda), colour=factor(bio_rep), h2o2, genotype, xlab="CFU", ylab="GDA (µg/ml)")) +
    scale_x_continuous(breaks = c(25,50,100,200,400,800)) +
    coord_trans('log10')

ggsave(ffFigFolder("oxidative_stress.png"), oxidative_stress_plot, height=6)
```

# Manduca assays
survival curves, grouped by kinase group and compared to wild type

```{r survival}
manduca_metadata <- read_excel(ffDataFolder("manduca_07march17.xlsx"), sheet="code")
manduca_survival <- list.files(path = ffDataFolder(), pattern = "^manduca_.*.xlsx") %>% 
  ffDataFolder() %>% 
  lapply(read_excel) %>%
  bind_rows() %>% 
  left_join(tibble(group=c(rep('ck2 complex', 4), 
                                rep('fungal-specific', 4), 
                                rep('conserved', 4)), 
                        treatment=c('cka1', 'cka2', 'ckb1', 'ckb2',
                                    'fun31', 'thi20', 'rim15', 'rim11',
                                    'hog1', 'mkk2', 'pkc1', 'pbs2')))

make_surv_plot <- function(data_t) {
 fit1 <- survfit(Surv(data_t$futime, data_t$fustat) ~ treatment, data=data_t) 
 labels <- str_remove(names(fit1$strata), 'treatment=')
 ggsurvplot(fit1, legend='bottom',
            legend.labs=labels, censor=FALSE)$plot + theme(legend.title=element_blank())
}

data_t <- manduca_survival %>% 
  select(-weight) %>% 
  pivot_wider(names_from=hours_pi, values_from=alive) %>% 
  mutate(futime=as.integer(case_when(`24` == 0 ~ 0,
                                     `48` == 0 ~ 1,
                                     `72` == 0 ~ 2,
                                      TRUE ~ 3)),
         fustat=as.integer(ifelse(is.na(`72`) | `72`==0, 1, 0)))

(conserved_plot <- data_t %>% filter(group == 'conserved' | treatment == 'wt' | treatment =='pbs') %>% 
  make_surv_plot() + scale_colour_brewer(palette = "Set1"))

ggsave(ffFigFolder("conserved_survival.png"), conserved_plot)

(fungal_specific_plot <- data_t %>% filter(group == 'fungal-specific' | treatment == 'wt' | treatment =='pbs') %>% 
  make_surv_plot() + scale_colour_brewer(palette = "Set1"))

ggsave(ffFigFolder("fungal_specific_survival.png"), fungal_specific_plot)

(ck2_plot <- data_t %>% filter(group == 'ck2 complex' | treatment == 'wt' | treatment =='pbs') %>% 
  make_surv_plot() + scale_colour_brewer(palette = "Set1"))

ggsave(ffFigFolder("ck2_survival.png"), ck2_plot)
```
