---
title: "Hsp90 Phosphorylation Blocks Fungal Virulence"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(folderfun)
library(pzfx)
library(ggthemes)


setff("DataFolder", file.path("T25", "data_files"))
setff("FigFolder", file.path("T25", "Figures"))
dir.create(ffFigFolder(), showWarnings = FALSE)

genotypes=data.frame(YSD=c("89", "1384", "1403", "1410", "1442", 
                           "1444", "557", "622", "628", "634"), 
                     genotype=factor(c("WT", "Mal2-Hsp90", "T25A", "T25E", "S530A", 
                                       "S530E", "cka1∆/∆", "cka2∆/∆", "ckb1∆/∆", "ckb2∆/∆"),
                                     levels=c("WT", "Mal2-Hsp90", "T25A", "T25E", "S530A",
                                              "S530E", "cka1∆/∆", "cka2∆/∆", "ckb1∆/∆", "ckb2∆/∆")))

```

# Fig 5b: Biofilm cell viability
- Plotted fluorescence per strain, with mean and standard error
- I also fit a linear linear model of fluorescence by strain, using Mal2-Hsp90 as the reference
- 530E had significantly less fluorescence than Mal2-Hsp90. WT was not signifincatly different. All other strains had significantly higher fluorecence
```{r cell_viability}
Data3 <- read_pzfx(ffDataFolder("CalbicansMutantsCV_REZ.pzfx"),
                   table='Data 3') %>% 
  gather(strain, fluorescence) %>% 
  left_join(genotypes, by=c("strain" = "YSD")) 

(bf_fluorescence <- Data3 %>% ggplot(aes(x=genotype, y=fluorescence)) +
    geom_jitter(width=.2, alpha=.3) +
    stat_summary(fun.data = mean_se, geom = "errorbar", width=.2) +
    stat_summary(fun.y = mean, geom = "point", size=2, colour='red') +
    theme_bw()
  )

ggsave(ffFigFolder("bf_fluorescence.pdf"), bf_fluorescence)

lm(fluorescence ~ genotype, data=mutate(Data3, genotype=relevel(genotype, ref="Mal2-Hsp90"))) %>% summary()
```

# Fig S3: Phosphorylation of S530 releases Hsp90’s block on the yeast-to-hyphae transition
- Plotted morphology index for each strain when grown on maltose and dextrose
- I also ran Mann-Whitney tests on maltose vs dextrose for each strain
- Morphology index higher in dextrose for T25A, T25E and S530E
- HIgher in maltose for Mal2-Hsp90 and no difference for WT and S530A
```{r}
morph_index <- read_delim(ffDataFolder("morph_index.txt"),
"\t", escape_double = FALSE, trim_ws = TRUE)

morph_index <- morph_index %>% 
  mutate(YSD=as.character(YSD)) %>% 
  full_join(genotypes) 

morph_index %>% 
  filter(!genotype %in% c("cka1∆/∆", "cka2∆/∆", "ckb1∆/∆", "ckb2∆/∆")) %>%
  ggplot(aes(x=Medium, y=MI, color=Medium)) +
    facet_grid(.~genotype) +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          legend.position = "bottom",
          strip.text=element_text(size=7)) +
    geom_jitter(size = .5) +
    scale_y_log10() +
    scale_color_manual(values=c(calc_pal()(12)[2],
                                calc_pal()(12)[1]))

ggsave(ffFigFolder("morphology.pdf"), bg="transparent")

# MI is not normally distributed so we use Man-Whitney U instead to t-test to compare two samples (maltose v. dextrose)
# non-parametric ANOVA is hard, so we are not comparing WT vs. mutants for each condition.

wilcox.test(filter(morph_index, genotype=='WT' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='WT' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='Mal2-Hsp90' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='Mal2-Hsp90' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='T25A' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='T25A' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='T25E' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='T25E' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='S530A' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='S530A' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='S530E' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='S530E' & Medium=='maltose') %>% `$`(MI))
```




