---
title: "Characterization of a novel Hsp90 phosphoswitch"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(folderfun)
library(pzfx)
library(ggthemes)


setff("DataFolder", file.path(Sys.getenv("DiezmannLab"), "T25", "data_files"))
setff("FigFolder", file.path(Sys.getenv("DiezmannLab"), "T25", "Figures"))
```

# Fig 5b: Biofilm cell viability
- Plotted fluorescence per strain, with mean and standard error
- I also fit a linear linear model of fluorescence by strain, using Mal2-Hsp90 as the reference
- 530E had significantly less fluorescence than Mal2-Hsp90. WT was not signifincatly different. All other strains had significantly higher fluorecence
```{r cell_viability}
Data3 <- read_pzfx(ffDataFolder("CalbicansMutantsCV_REZ.pzfx"),
                   table='Data 3') %>% 
  gather(strain, fluorescence) %>%
    mutate(strain=factor(strain, 
                         levels=c('89', '1384', '1403', '1410', 
                                  '1442', '1444')))

(bf_fluorescence <- Data3 %>% ggplot(aes(x=strain, y=fluorescence)) +
    geom_jitter(width=.2, alpha=.3) +
    stat_summary(fun.data = mean_se, geom = "errorbar", width=.2) +
    stat_summary(fun.y = mean, geom = "point", size=2, colour='red') +
    theme_bw()
  )

ggsave(ffFigFolder("bf_fluorescence.pdf"), bf_fluorescence)

lm(fluorescence ~ strain, data=mutate(Data3, strain=relevel(strain, ref='1384'))) %>% summary()
```

# Fig 5c: Metabolic biomass
- Plotted OD570 per strain, with mean and standard error
- I also fit a linear linear model of OD570 by strain, using Mal2-Hsp90 as the reference
- WT and S530A had significantly higher OD570 than Mal2-Hsp90. All other strains were not significantly different
```{r}
Data1 <- read_pzfx(ffDataFolder("CalbicansMutantsCV_REZ.pzfx"), table='Data 1') %>%
  gather(strain, OD570) %>%
  mutate(strain=factor(strain, 
                       levels=c('89', '1384', '1403',
                                '1410', '1442', '1444')))

(bf_OD570 <- Data1 %>%
  ggplot(aes(x=strain, y=OD570)) +
    geom_jitter(width=.2, alpha=.3) +
    stat_summary(fun.data = mean_se, geom = "errorbar", width=.2) +
    stat_summary(fun.y = mean, geom = "point", size=2, colour='red') +
    theme_bw()
  )
ggsave(ffFigFolder("bf_OD570.pdf"), bf_OD570)

lm(OD570 ~ strain, data=mutate(Data1, strain=relevel(strain, ref='1384'))) %>% summary()
```

# Fig S2: Yeast to hyphae transition
- Plotted morphology index for each strain when grown on maltose and dextrose
- I also ran Mann-Whitney tests on maltose vs dextrose for each strain
- Morphology index higher in dextrose for T25A, T25E and S530E
- HIgher in maltose for Mal2-Hsp90 and no difference for WT and S530A
```{r}
morph_index <- read_delim(ffDataFolder("/morph_index.txt"),
"\t", escape_double = FALSE, trim_ws = TRUE)

genotypes=data.frame(YSD=c("89", "1384", "1403", "1410", "1442", 
                           "1444", "557", "622", "628", "634"), 
                     genotype=factor(c("WT", "Mal2-Hsp90", "T25A", "T25E", "S530A", 
                                       "S530E", "cka1∆/∆", "cka2∆/∆", "ckb1∆/∆", "ckb2∆/∆"),
                                     levels=c("WT", "Mal2-Hsp90", "T25A", "T25E", "S530A",
                                              "S530E", "cka1∆/∆", "cka2∆/∆", "ckb1∆/∆", "ckb2∆/∆")))

morph_index <- morph_index %>% 
  mutate(YSD=as.character(YSD)) %>% 
  full_join(genotypes) 

morph_index %>% 
  filter(!genotype %in% c("cka1∆/∆", "cka2∆/∆", "ckb1∆/∆", "ckb2∆/∆")) %>%
  ggplot(aes(x=Medium, y=MI, color=Medium)) +
    facet_grid(.~genotype) +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          legend.position = "bottom",
          strip.text=element_text(size=7)) +
    geom_jitter(size = .5) +
    scale_y_log10() +
    scale_color_manual(values=c(calc_pal()(12)[2],
                                calc_pal()(12)[1]))

ggsave(ffFigFolder("morphology.pdf"), bg="transparent")

# MI is not normally distributed so we use Man-Whitney U instead to t-test to compare two samples (maltose v. dextrose)
# non-parametric ANOVA is hard, so we are not comparing WT vs. mutants for each condition.

wilcox.test(filter(morph_index, genotype=='WT' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='WT' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='Mal2-Hsp90' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='Mal2-Hsp90' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='T25A' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='T25A' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='T25E' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='T25E' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='S530A' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='S530A' & Medium=='maltose') %>% `$`(MI))


wilcox.test(filter(morph_index, genotype=='S530E' & Medium=='dextrose') %>% `$`(MI),
            filter(morph_index, genotype=='S530E' & Medium=='maltose') %>% `$`(MI))

```

# Tbl S6: Manduca survival
- Significant differences in the number of surviving manduca on day 1 by fisher's exact test
- Relative to manduca infected with MAL2-HSP90, manduca infected with 530E and with mock infection (PBS) had significantly higher survival
```{r survival}
manduca_survival <- read_tsv(ffDataFolder("manduca_survival.txt")) %>%
  rename(Strain=Treatment) %>%
  mutate(Strain=factor(Strain, levels=c('PBS', 'YSD89', 'YSD1384', 'YSD1403', 'YSD1410', 'YSD1442', 'YSD1444')))

manduca_survival %>% 
  select(Strain, Day2_AM) %>%
  mutate(Strain=as.character(Strain)) %>%
  table() %>%
  fisher.test()

strain_names <- tribble(~strain, ~genotype,
                        "PBS",     "PBS",
                        "YSD89",   "WT",
                        "YSD1384", "MAL2-HSP90",
                        "YSD1403", "T25A",
                        "YSD1442", "T25E",
                        "YSD1442", "S530A",
                        "YSD1444", "530E")
strain <- "YSD1403" 

run_fisher_test <- function(strain, df) {
  df %>% 
  select(Strain, Day2_AM) %>%
  mutate(Strain=as.character(Strain)) %>%
  filter(Strain %in% c("YSD1384", strain)) %>%
  table() %>%
  fisher.test() %>% 
  `$`(p.value) 
}

strain_names$pval <- map_dbl(strain_names$strain, possibly(run_fisher_test, NA), manduca_survival)
strain_names
```



